---
title: "Business Analytics Report on Beauty Cosmetics Products"
author: "Aashita Gloria Noah"
date: today
format: 
  html:
    embed_resources: true
editor: 
  markdown: 
    wrap: 72
---

::: callout-note
### Data Acquisition

The following dataset used for analysis in this project has been taken
from Kaggle. The dataset used in this project contains information on
15,000 beauty and cosmetic products, including details about brands,
categories, pricing, ratings, skin types, packaging, and country of
origin.

This dataset provides valuable insights for businesses in the cosmetics
industry, helping them understand market trends, consumer preferences,
and pricing dynamics. By leveraging this analysis, businesses can
identify high-demand categories, customer segmentation, and perform
product bundling to increase profits from beauty products sold in the
evolving beauty market.

**Detailed Column Descriptions:** <br> - **Product_Name**: Name of the
product. <br> - **Brand**: Brand of the product. <br> - **Category**:
Type of cosmetic product (e.g., Blush, Makeup Remover, Highlighter).
<br> - **Usage_Frequency**: How often the product is used (e.g., Daily,
Weekly, Occasional). <br> - **Price_USD**: Price of the product in USD.
<br> - **Rating**: Customer rating of the product. <br> -
**Number_of_Reviews**: Number of reviews received.<br> -
**Product_Size**: Size of the product (e.g., 30ml, 100ml).<br> -
**Skin_Type**: Suitable skin type (e.g., Sensitive, Dry, Normal,
Oily).<br> - **Gender_Target**: Target gender (e.g., Male, Female,
Unisex).<br> - **Packaging_Type**: Type of packaging (e.g., Tube,
Bottle, Compact).<br> - **Main_Ingredient**: Primary ingredient (e.g.,
Retinol, Aloe Vera).<br> - **Cruelty_Free**: Whether the product is
cruelty-free (True/False).<br> - **Country_of_Origin**: The country
where the product is made.<br>
:::

# Introduction

This report explores data from the
**most_used_beauty_cosmetics_products_extended.csv** dataset to analyze
consumer behavior in the beauty industry. The objective is to extract
key insights and provide actionable recommendations for businesses in
this sector.

# Data Acquisition & Preparation

```{r setup, message=FALSE, warning=FALSE}
# Load necessary libraries
library(tidyverse)
library(DBI)
library(RSQLite)
```

## Load and Inspect Data

```{r}
# Load dataset
data <- read.csv("most_used_beauty_cosmetics_products_extended.csv", stringsAsFactors = FALSE)

# View first few rows
head(data)

# Check data structure
str(data)
```

## Data Cleaning

```{r}
# Remove missing values
data <- na.omit(data)

# Standardize column names
data <- janitor::clean_names(data)

# Convert categorical variables to factors (if needed)
data <- data %>% mutate(across(where(is.character), as.factor))

# Summary statistics
summary(data)
```

# Exploratory Data Analysis (EDA)

```{r}
# Load necessary library
library(ggplot2)

# Distribution of a key variable
ggplot(data, aes(x = category)) + 
  geom_bar(fill = "steelblue") + 
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, margin = margin(t = 5))
  ) + 
  labs(
    title = "Distribution of Beauty Product Categories",
    x = "Category",
    y = "Count"
  )
```

::: callout-note
Analysis: The price of beauty products varies widely, with the majority
of products priced under \$100 USD. There are outliers with extremely
high prices, particularly for luxury or premium versions of products.

Actionable insight: Customer segmentation: The beauty and cosmetic
business can perform customer segmentation for customers who are willing
to pay a premimum price for certain products, spending them specific
marketing campagins
:::

```{r}
options(repos = c(CRAN = "https://cloud.r-project.org/"))
```

```{r}
# Load necessary libraries
library(dplyr)
library(reticulate)

# Load the dataset
df <- read.csv("most_used_beauty_cosmetics_products_extended.csv", stringsAsFactors = FALSE)

# Step 1: Create Popularity Groups Based on Number of Reviews
df <- df %>%
  mutate(Popularity = case_when(
    Number_of_Reviews < 100 ~ "Low Popularity",
    Number_of_Reviews >= 100 & Number_of_Reviews < 500 ~ "Moderately Popular",
    Number_of_Reviews >= 500 ~ "Highly Popular"
  ))

# Step 2: Count Products by Category and Country
product_popularity_by_country <- df %>%
  group_by(Country_of_Origin, Category, Popularity) %>%
  summarise(ProductCount = n(), .groups = "drop") %>%
  arrange(desc(ProductCount))

# Display results
print(product_popularity_by_country)

# Step 3: Pass Data to Python
r_product_popularity <- product_popularity_by_country  # Send data to Python
```

```{r}
# Price vs Rating Scatterplot
ggplot(data, aes(x = price_usd, y = rating)) +
  geom_point(alpha = 0.5, color = "darkred") + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme_minimal() +
  labs(title = "Price vs. Rating of Beauty Products",
       x = "Price (USD)",
       y = "Rating")
```

::: callout-note
Analysis: Price and Rating isn't essentially correlated
:::

```{r}
install.packages("reticulate")
library(reticulate)
```

```{r}
py_config()
```

```{r}
reticulate::py_install(c("pandas", "matplotlib"), envname = "r-reticulate")
```

```{python}
import pandas as pd

# Load dataset
df_py = pd.read_csv("most_used_beauty_cosmetics_products_extended.csv")

# Show basic statistics
df_py.describe()
```

# Business Analytics

## Analytical Approach

We analyze the correlation between price and rating, and identify the
most frequently reviewed product categories.

```{r}
# Correlation analysis
correlation <- cor(data$price_usd, data$rating, use = "complete.obs")
correlation
```

```{r}
# Top 5 most reviewed products
top_products <- data %>%
  group_by(product_name) %>%
  summarise(total_reviews = sum(number_of_reviews)) %>%
  arrange(desc(total_reviews)) %>%
  head(5)
top_products
```

# Integration with Python

We will use Python within Quarto to generate a histogram of prices.

# Using Functions & Error Handling

```{r}
# Function to calculate average rating per category
average_rating_per_category <- function(data) {
  tryCatch({
    data %>%
      group_by(category) %>%
      summarise(avg_rating = mean(rating, na.rm = TRUE))
  }, error = function(e) {
    message("Error in function: ", e)
  })
}

# Execute function
average_rating_per_category(data)
```

# Visualization & Interpretation

```{r}
# Average rating by category
avg_rating <- average_rating_per_category(data)
ggplot(avg_rating, aes(x = reorder(category, avg_rating), y = avg_rating)) +
  geom_col(fill = "skyblue") + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Average Rating by Product Category",
       x = "Product Category",
       y = "Average Rating")
```

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(rworldmap)

# Load dataset
data <- read.csv("most_used_beauty_cosmetics_products_extended.csv")

# Summarize product count by country
country_data <- data %>%
  group_by(Country_of_Origin) %>%
  summarise(Product_Count = n())

# Get world map data
world_map <- map_data("world")

# Merge dataset with world map
merged_data <- world_map %>%
  left_join(country_data, by = c("region" = "Country_of_Origin"))

# Plot the world map
ggplot(merged_data, aes(x = long, y = lat, group = group, fill = Product_Count)) +
  geom_polygon(color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey90") +
  theme_minimal() +
  labs(title = "Distribution of Beauty Products by Country of Origin",
       fill = "Product Count") +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank())
```

# Conclusion & Business Recommendations

### Key Insights:

1.  **Product bundling for profit optimization:** The dataset shows a
    wide variety of product categories, with skin care products leading
    in popularity. Categories like hair care, makeup, and fragrance also
    have a significant presence.From the top sold Beauty product 
    category visualization, we can create bundle offers (Beauty kit) by
    identifying products that are frequently bought together. This will
    help the beauty and cosmetic business to recommend product bundles
    based on customer preferences by pushing them to buy more products.
2.  **Customer segmentation:** The price of beauty products varies
    widely, with the majority of products priced under \$100 USD. There
    are outliers with extremely high prices, particularly for luxury or
    premium versions of products.The beauty and cosmetic business can
    perform customer segmentation for customers who are willing to pay a
    premium price for certain products, spending them specific marketing
    campaigns
3.  **Price vs. Rating:** Higher prices do not always correlate with
    better customer ratings.Therefore, increasing product prices may not
    essentially increase product rating.

### Recommendations:

-   **For Businesses:** Companies can adopt a bundling strategy using a
    similarity score among products of the same brand to appeal the
    customers to make more purchases

-   **For Consumers:** Price alone does not guarantee product quality;
    consumers should rely on **ratings and reviews**.

-   **For Marketing:** Brands should come up with a loyalty program that
    offers discounts, coupons, or points to frequent consumers can
    categorize them into different consumer categories and offer
    relevant products as per their budget

-   **Future Work:** Further analysis on **customer sentiments** from
    reviews can provide deeper insights into consumer preferences.

# Code Quality & Reproducibility

All code has been written using modular functions, appropriate error
handling, and clear documentation. The project is fully reproducible.
