---
title: "firstQuartoDocument"
subtitle: "I am learning Quarto"
author: "Aashita Gloria Noah"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 4
editor: visual
execute:
  warning: false 
  message: false
  echo: false
---

## Loading libraries

```{r}

library(tidyverse)
library(gt)
library(knitr)
```

## Quarto

## Intro

This is the intro to our quarto document

***Quarto*** enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

## Comments

### Note

::: callout-note
This is a note
:::

### Important

::: callout-important
This is important
:::

### Warning

::: callout-warning
This is a warning
:::

### Tabsets

::: panel-tabset
## Using only head

```{r}
head(iris)
```

## Using nothing

```{r}
airquality
```

## datatable

```{r}
library(DT)
datatable(airquality)
```

## kable

```{r}
library(knitr)
kable(head(airquality))
```

## gt

```{r}
library(gt)
airquality |> 
  head() |>
  gt()
```

## tibble

```{r}
airquality |> 
  head() |>
  tibble()
```

:::

### 3 Column Solution

::: columns
:::: {.column width="45%"}

```{r}
head(iris)
```

::::
:::: {.column width="10%"}
::::

:::: {.column width="45%"}

```{r}
head(mtcars)
```

::::
:::
